Manuscript progress report

### 2024-02-07 Docs/NhanesApplExample.Rmd
try to simulate using estimated from debias model

I'd like to investigate with different sample sizes.

```{r, cache=TRUE, results='hide'}
# sample size 
N_sub <- rep(500, 4)
fpca_model_list <- list()
interp_pc_est <- list()
debias_model_list <- list()
time_vec <- rep(NA, length(N_sub))

pb <- txtProgressBar(min=0, max = length(N_sub), style = 3)
for(i in seq_along(N_sub)){

  # training sample of a certain size
  Ntr <- N_sub[i]
  train_id <- sample(unique(df$id), size = Ntr) 
  train_df <- df %>% filter(id %in% train_id)

@@

  # FPCA
  uni_eta_hat <- df_est_latent %>% filter(bin==sind)
  mat_est_unique <- matrix(uni_eta_hat$eta_hat, nrow=Ntr, ncol=n_bin, byrow = F) 
  fpca_mod <- fpca.face(mat_est_unique, argvals = mid, var=T, npc=K)
  fpca_model_list[[i]] <- fpca_mod

@@

  df_phi <- matrix(NA, J, K) 
  for(k in 1:K){
    lm_mod <- lm(fpca_mod$efunctions[,k] ~ B-1)
    df_phi[,k] <- Bnew %*% coef(lm_mod)
  }# project binned eigenfunctions onto the original grid
  interp_pc_est[[i]] <- df_phi

@@

  setTxtProgressBar(pb, i)
}

close(pb)
```

@@

>>>>

```{r, cache=TRUE, results='hide'}
# sample size 
Ntr <- 500

# training sample of a certain size
train_id <- sample(unique(df$id), size = Ntr) 
train_df <- df %>% filter(id %in% train_id)

@@

# FPCA
uni_eta_hat <- df_est_latent %>% filter(bin==sind)
mat_est_unique <- matrix(uni_eta_hat$eta_hat, nrow=Ntr, ncol=n_bin, byrow = F) 
fpca_mod <- fpca.face(mat_est_unique, argvals = mid, var=T, npc=K)

@@

df_phi <- matrix(NA, J, K) 
for(k in 1:K){
  lm_mod <- lm(fpca_mod$efunctions[,k] ~ B-1)
  df_phi[,k] <- Bnew %*% coef(lm_mod)
}# project binned eigenfunctions onto the original grid

@@

```

@@

```{r, class.source='fold-show'}
load(here("Data/Appl_debias_model.RData"))
summary(debias_glmm)

# re-evaluated mean
new_mu <- predict(debias_glmm, type = "terms")[1:J, 1] # extract re-evaluated mean
plot(1:J, new_mu)

# re-evaluated eigenvalues
new_lambda <- 1/debias_glmm$sp[2:5]
new_lambda <- new_lambda/n_bin

# re-evaluated PC functions
df_phi <- debias_glmm$model %>% select(sind, starts_with("phi")) %>% distinct(.) %>%
  mutate_at(vars(starts_with("phi")), function(x){x*sqrt(n_bin)})
df_phi %>% pivot_longer(2:5) %>%
  ggplot()+
  geom_line(aes(x=sind, y=value, col = name))

```

```{r gen_data}
# generate data with the estimates from thedebias step

# generate score
xi <- rmvnorm(Ntr, mean = rep(0, K), sigma = diag(new_lambda))
bi <- xi %*% t(as.matrix(df_phi[, 2:5])) # random effect functions
dim(bi)

# latent gaussian function
new_mu <- matrix(new_mu, nrow = 1)
eta_i <- new_mu+bi
dim(new_mu)

  this_df <- data.frame(id = factor(rep(1:N, each=J)),
                        t = rep(t, N), 
                        eta_i = as.vector(t(eta_i)),
                        sind = rep(1:J, N))

  this_df$Y <- rbinom(N*J, size=1, prob=plogis(this_df$eta_i))

  sim_data[[m]] <- this_df
]
```

### 2024-02-07 Docs/manuscript.Rmd
try to simulate using estimated from debias model

pred_appl_gfosr_l1 <- pred_appl_gfosr_l1 %>%
  rename(pred360=pred_w1, pred720=pred_w2, pred1080=pred_w3)
pred_appl_gfosr_l5 <- pred_appl_gfosr_l5 %>%
  rename(pred360=pred_w1, pred720=pred_w2, pred1080=pred_w3)

@@

```{r fig_appl, fig.height=8, fig.width=8}
rand_id <- sample(unique(pred_nhanes_adglmm$id), 4)
cols <- cols[1:3]
names(cols) <- c("6am", "12pm", "6pm")
bind_rows(
  pred_nhanes_fgfpca %>% filter(id %in% rand_id) %>% mutate(method="fGFPCA"),
  pred_appl_gfosr_l5 %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=5)"),
  pred_appl_gfosr_l1 %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=1)"),
  pred_nhanes_adglmm %>% filter(id %in% rand_id) %>% mutate(method = "GLMMadaptive")) %>%
  mutate_at(vars(pred360, pred720, pred1080), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive"))) %>%
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.2)+
  geom_line(aes(x=sind, y=pred360, col = "6am"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=sind, y=pred720, col = "12pm"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=sind, y=pred1080, col = "6pm"), linetype="dashed", na.rm=T)+
  facet_grid(rows = vars(id), cols = vars(method))+
  labs(col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1440, by = 360),
                     labels = c("0am", "6am", "12pm", "6pm", "12am"))+
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7))+
  scale_color_manual(values = cols)
```

@@

>>>>

pred_appl_gfosr_l1 <- pred_appl_gfosr_l1 %>%
  rename(pred360=pred_w1, pred720=pred_w2, pred1080=pred_w3)
pred_appl_gfosr_l5 <- pred_appl_gfosr_l5 %>%
  rename(pred360=pred_w1, pred720=pred_w2, pred1080=pred_w3)
```

@@

```{r fig_appl, fig.height=8, fig.width=8}
rand_id 
cols <- cols[1:3]
names(cols) <- c("6am", "12pm", "6pm")
bind_rows(
  pred_nhanes_fgfpca %>% filter(id %in% rand_id) %>% mutate(method="fGFPCA"),
  pred_appl_gfosr_l5 %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=5)"),
  pred_appl_gfosr_l1 %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=1)"),
  pred_nhanes_adglmm %>% filter(id %in% rand_id) %>% mutate(method = "GLMMadaptive")) %>%
  mutate_at(vars(pred360, pred720, pred1080), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive"))) %>%
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.2)+
  geom_line(aes(x=sind, y=pred360, col = "6am"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=sind, y=pred720, col = "12pm"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=sind, y=pred1080, col = "6pm"), linetype="dashed", na.rm=T)+
  facet_grid(rows = vars(id), cols = vars(method))+
  labs(col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1440, by = 360),
                     labels = c("0am", "6am", "12pm", "6pm", "12am"))+
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7))+
  scale_color_manual(values = cols)
```

@@

```{r eval=FALSE}
df_nhanes <- read_rds(here("Data/nhanes_bi.rds"))
df_nhanes %>% 
  select(SEQN, Z, sind) %>%
  filter(SEQN %in% rand_id) %>%
  filter(sind %in% 356:360) %>%
  pivot_wider(names_from = sind, values_from = Z) %>% 
  select(-SEQN) %>% distinct(.)

pred_appl_gfosr_l5 %>% 
  filter(id %in% rand_id) %>%
  select(id, sind, pred360) %>%
  pivot_wider(names_from = sind, values_from = pred360) %>% 
  select(-id) %>% distinct(.)
```
